---
title: "Clinical Trial Practical"
author: "Sachin"
date: "27/01/2023"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

# 1) Randomization Methods
# Complete Randomization
```{r}
x=c(1,2,3,4)
p=c(0.2,0.4,0.3,0.1)
drug=sample(x,100,replace=T,prob=p)
d1=data.frame('Subject'=seq(1,100),drug)
tx=table(drug);tx
```
# Permuted Block Randomization
```{r}
n=30
blocks=3
bsize=n/blocks
drug=c(rep('T',bsize/2),rep('R',bsize/2))
t1=sample(drug,replace = F)
t2=sample(drug,replace=F)
t3=sample(drug,replace = F)
t=c(t1,t2,t3)
d2=data.frame('Patient'=seq(1:n),"Treatment"=t)
d3=table(t);d3
```
# Adaptive Randomization
# 1) Treatment Adaptive Randomization
```{r}
tre=rep(0,10)
drug=c("T","R")
p=.5
s=.02
tre[1]=sample(drug,1,prob=c(p,1-p))
for(i in 2:10){
  p=ifelse('T'==tre[i-1],p-s,p+s)
  tre[i]=sample(drug,1,prob=c(p,1-p))
}
data.frame("Subject"=seq(1:10),"Drug"=tre)
```
# 2) Response Adaptive Randomization
```{r}
tre=rep(0,10)
res=rep(0,10)
drug=c('T',"R")
p=.5
s=.02
q=.6
tre[1]=rbinom(1,1,p)
res[1]=rbinom(1,1,q)
for(i in 2:10){
  p=ifelse(tre[i-1]==res[i-1],p+s,p-s)
  tre[i]=rbinom(1,1,p)
  res[i]=rbinom(1,1,q)
}
data.frame("Subject"=seq(1:10),"Drug"=drug[tre+1])
```
# 3) Covariate Adaptive Randomization
```{r}
sub=c(1:20)
gender=c('M','M','F','M','F','F','F','M','M','F','F','M','M',
         'M','F','M','F','M','M','M')
smoke=c('No','Yes','Yes','No','No','Yes','No','No','No','Yes',
        'Yes','Yes','No','Yes','No','Yes','No','Yes','Yes','No')
FS=sub[gender=='F'&smoke=='Yes']
nFS=length(FS)
Y1=sample(c(rep(c('T',"R"),nFS/2)),replace = F)
FN=sub[gender=='F'& smoke=='No']
nFN=length(FN);Y2=sample(c(rep(c('T','R'),nFN/2)),replace=F)
MS=sub[gender=='M'& smoke=='Yes']
nMS=length(MS);Y3=sample(c(rep(c('T','R'),nMS/2)),replace=F)
MN=sub[gender=='M'& smoke=='No'];
nMN=length(MN);Y4=sample(c(rep(c('T','R'),nMN/2)),replace=F)
da=data.frame(subject=c(FS,FN,MS,MN),drug=c(Y1,Y2,Y3,Y4))
da
```
# 2)  Treatment Comparisons in Clinical Trials
# 1)  Analysis of Parallel Design
```{r}
x=c(2.5,2.25,2.35,2.60,2.10,2.45,2.50,2.1,2.2)
y=c(2.45,2.50,2.60,2.77,2.60,2.55,2.65,2.75,2.45,2.50)
t.test(x,y,alt="less",var.equal=TRUE)
```
# For two sample assuming unequal variances
```{r}
t.test(x,y,alt="less",var.equal=FALSE) # use `paired = T' for paired t-test
```
# For one sample:
# A sample 12 student from a school has the following scores in an IQ. test. 89, 87, 76, 78, 79, 86, 74,
# 83, 75, 71, 76 and 92. Do this data support that the mean IQ marks of the school students is 80. Test
# at 5%level
```{r}
x=c(89,87,76,78,79,86,74,83,75,71,76,92)
t.test(x,mu=80,alt="two.sided",conf.level = 0.95)
```
# 2) Categorical Endpoints: Pearson's chisquare-test
```{r}
x = matrix(c(45,56,35,47),ncol=2)
prop.test(x,alternative = "two.sided")
```
# 3) Analysis of Variance
```{r}
y1=c(12,12.3,15,14.5,15.2,13.6,14.6,12.5,12.4);n1=length(y1);
y2=c(24.2,26.3,20.4,21.5,24.3,22,25.6,23.5);n2=length(y2);
y3=c(24.3,22,24,25.1,21.4,20,12);n3=length(y3);
y4=c(30.4,29.4,28.4,25.45,27.12,20.4,28,29.2,30.4);n4=length(y4);
yield=c(y1,y2,y3,y4);
fact=factor(c(rep(1,n1),rep(2,n2),rep(3,n3),rep(4,n4)));
model1<-aov(yield~fact);
summary(model1)
```
# Pk Parameter
```{r}
data=read.csv("C:/MSC-3 Sem/ST-306/B1/MCP/BABEData.csv",header=F)
data=data[,5:20]
ti=unlist(data[1,])
ci=unlist(data[2,])

n=length(ci)
cmax=max(ci)
a=which.max(ci)
tmax=ti[a]
auct=sum(ci[2:n]+ci[1:n-1]*ti[2:n]-ti[1:n-1])/2
ke=-2.303*coefficients(lm(log10(ci[(a+1):n])~ti[(a+1):n]))[2]    # Rate of elimination
thalf=0.693/ke
aucinf=auct+ci[n]/ke
df=data.frame("Cmax"=cmax,"tmax"=tmax,"auct"=auct,'Ke'=ke,"AUCinf"=aucinf,"Thalf"=thalf)

pkpara=function(ci,ti){
n=length(ci)
cmax=max(ci)
a=which.max(ci)
tmax=ti[a]
auct=sum((ci[2:n]+ci[1:n-1]*ti[2:n]-ti[1:n-1]))/2
ke=-2.303*coefficients(lm(log10(ci[(a+1):n])~ti[(a+1):n]))[2]
thalf=0.693/ke
aucinf=auct+ci[n]/ke
df=data.frame("Cmax"=cmax,"tmax"=tmax,"auct"=auct,'Ke'=ke,"AUCinf"=aucinf,"Thalf"=thalf)
return(df)
}
pkpara(ci,ti)
df=data.frame("Cmax"=0,"tmax"=0,"auct"=0,"Thalf"=0,'Ke'=0,"AUCinf"=0)
for(i in 2:49){
  df=rbind(df,pkpara(unlist(data[i,]), unlist(data[1,])))
}
df
```
# Shortest confidence interval
```{r}
data=read.csv('C:/MSC-3 Sem/ST-306/B1/MCP/BABEData22.csv')
y11=data[data$Seq==1&data$Per==1,1];n1=length(y11);
y12=data[data$Seq==1&data$Per==2,1];n2=length(y12);
y21=data[data$Seq==2&data$Per==1,1];n3=length(y21);
y22=data[data$Seq==2&data$Per==2,1];n4=length(y22);
y1=c(y11,y12)
y2=c(y21,y22)
y1bar=mean(y1)
y2bar=mean(y2)
c=y2bar-y1bar
s1=var(y1)
s2=var(y2)
sigma2=((n1-1)*s1+(n2-1)*s2)/(n1+n2-2)
sd=sqrt(sigma2)
z=qnorm(1-0.05/2)
ci=c(c-z*sd*sqrt(1/n1+1/n2),c+z*sd*sqrt(1/n1+1/n2))
ci
```
# Sample size determination
```{r}
alpha = 0.05; 
beta = c(0.05,0.1,0.15,0.2,0.25,0.3); 
delta=0.5;
pow = 1-beta
# Required sample size
num = 2*(qnorm(1-alpha/2)+qnorm(1-beta))^2/delta^2;
num
ceiling(num)
```





